---
title: "Analysis of NMO party representation"
author: "Zoe MacLean"
date: '2023'
output:
  word_document: default
---
Code in this analysis was based on tutorial material made available by Dr Martin Schweinberger of the University of Queensland, available on the LADAL website.

Schweinberger, M. (2022, September 14). Fixed- and Mixed-Effects Regression Models in R. The University of Queensland. https://slcladal.github.io/regression.html

**Preparation and session set up**


```{r}
# load packages
library(ggplot2)
library(gridExtra)
library(glmulti)
library(lme4)
library(rms)
library(sjPlot)
library(here)
library(dplyr)
library(car)
library(mclogit)
library(broom)
library(tidyverse)
library(car)
library(emmeans)
```


Set options

```{r}
# set options
options("scipen" = 100, "digits" = 12) # suppress math annotation
```

# Valence Analysis

Load the data

```{r}
data <- read.delim(here::here("data", "text_valence.csv"), sep = ",", header = T)
```
Factorise the relevant character based variables. Do this one variable at a time to avoid postid. 

```{r}
# This factorises the organisation column (6 levels)
data$ownership <- as.factor(data$ownership)

# This factorises the title column (297 levels)
data$nmo_title <- as.factor(data$nmo_title)

# This factorises the party mention column (3 levels)
data$party_mentioned <- as.factor(data$party_mentioned)


```

Inspect the data

```{r}
summary(data)
```

```{r}
num_unique_postid <- length(unique(data$postid))
cat("Number of unique values in 'postid':", num_unique_postid)
```
There are 10,436 mentions in 8,723 unique facebook posts. Each post can mention one or more political parties, hence we have more mentions than posts in the dataset.

Create a two-way ANOVA model

```{r}
model <- aov(valence ~ ownership * party_mentioned, data = data)
```

Check the model's assumptions with diagnostic plots:

```{r}
# Open a PDF file for plotting
pdf(file = "diagnostic_plots.pdf", width = 14, height = 10)

# Set up a 2x2 grid for plotting
par(mfrow = c(2, 2))

# Residuals vs. Fitted values
plot(model, 1)

# Normal Q-Q plot
plot(model, 2)

# Scale-Location plot
plot(model, 3)

# Cook's distance plot
plot(model, 4)

# Close the PDF file
dev.off()

# Residuals vs. Fitted values
plot(model, 1)

# Normal Q-Q plot
plot(model, 2)

# Scale-Location plot
plot(model, 3)

# Cook's distance plot
plot(model, 4)

```

Check for homoscedasticity using the Levene's Test:

```{r}
leveneTest(valence ~ ownership * party_mentioned, data = data)
```

Create a histogram of the residuals:

```{r}
hist(resid(model), main = "", xlab = "Residuals")

# Open a PDF file for plotting
pdf(file = "histogram_resid.pdf", width = 10, height = 5)

# Plot the histogram
hist(resid(model), main = "", xlab = "Residuals")

# Close the PDF file
dev.off()

```

Create a histogram of the valence scores

```{r}
# Use a histogram to check the distribution of valence scored
hist(data$valence, main = "", xlab = "Valence")

# Open a PDF file for plotting
pdf(file = "histogram_val.pdf", width = 10, height = 5)

# Plot the histogram
hist(data$valence, main = "", xlab = "Valence")

# Close the PDF file
dev.off()

```


Run the Kruskal-Wallis test:

```{r}
data$group <- interaction(data$ownership, data$party_mentioned)
kruskal.test(valence ~ group, data = data)
```
Run the two-way ANOVA:

```{r}
anova_results <- anova(model)
print(anova_results)

```

Visualise the difference in valence scores:

```{r}
# Hex codes based on the colours of the party logos
party_colors <- c("alp" = "#d34547", "gre" = "#3b874a", "coa" = "#2a4e96") 

valence_scores<- ggplot(data, aes(x = ownership, y = valence, fill = party_mentioned)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "",
       x = "Organisation",
       y = "Valence Score",
       fill = "Political Party") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = party_colors,
                    labels = c("Labor", "The Coalition","The Greens"))

ggsave(filename = "valence_scores.pdf", plot = valence_scores, device = "pdf", width = 8, height = 5, dpi = 600)

valence_scores 
```



Now use a mixed effects multinomial logistic regression model.

```{r}
mixed_model <- mblogit(ownership ~ party_mentioned + valence, random = ~ 1 | nmo_title, data = data)
```


```{r}
summary(mixed_model)
```

Validate the use of random effects by creating a fixed-effects version of the model and then compare AIC and BIC scores.

```{r}
fixed_model <- mblogit(ownership ~ party_mentioned + valence, data = data)
```

```{r}
AIC_fixed <- AIC(fixed_model)
BIC_fixed <- BIC(fixed_model)

AIC_mixed <- AIC(mixed_model)
BIC_mixed <- BIC(mixed_model)

cat("AIC Fixed:", AIC_fixed, "\n")
cat("AIC Mixed:", AIC_mixed, "\n")
cat("BIC Fixed:", BIC_fixed, "\n")
cat("BIC Mixed:", BIC_mixed, "\n")
```
Use the `emmeans` package to perform pairwise comparisons of estimated marginal means. 

```{r}
emm_model <- emmeans(mixed_model, ~ ownership | party_mentioned, type = "response")
pairs(emm_model, adjust = "tukey")
```
Summary of the mixed-effects model:

```{r}
mclogit::getSummary.mmblogit(mixed_model)
```

Check the VIFs to see if the model does not violate multicollinearity assumptions.

```{r}
car::vif(mixed_model) 
```

# Image Analysis

Load the data

```{r}
data_img <- read.delim(here::here("data", "emotions_faces.csv"), sep = ",", header = T)
```
Factorise the relevant character based variables.

```{r}
# This factorises the organisation column (6 levels)
data_img$ownership <- as.factor(data_img$ownership)

# This factorises the title column (297 levels)
data_img$nmo_title <- as.factor(data_img$nmo_title)

# This factorises the party mention column (3 levels)
data_img$party_mentioned <- as.factor(data_img$party_mentioned)

# and this factorises the type of mention (2 levels)
data_img$dominant_emotion <- as.factor(data_img$dominant_emotion)
```

Inspect the data

```{r}
summary(data_img)
```

```{r}
num_unique_postid <- length(unique(data_img$postid))
cat("Number of unique values in 'postid':", num_unique_postid)
```

5,131 mentions in 4,236 unique posts.

```{r}
# Save plot to a variable
probability_own <- ggplot(data_img, aes(ownership, happy_bin, color = party_mentioned)) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(0.2)) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2, position = position_dodge(0.2)) +
  theme_set(theme_bw(base_size = 10)) +
  theme(legend.position = "right") +  # Change legend position to right
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  labs(
    title = "",
    x = "Ownership",
    y = "Observed Probability of Happy",
    color = "Political Party"
  ) +
  scale_color_manual(
    values = c("#d34547", "#2a4e96", "#3b874a"),
    labels = c("Labor", "The Coalition", "The Greens")
  )

# Save plot to PDF
ggsave(filename = "probability_own.pdf", plot = probability_own, width = 8, height = 5, dpi = 600)

probability_own

```


First, cut the data down to only emotions over the 0.80 threshold no matter the type of emotion
```{r}
# Subset the data to exclude any emotion scores below .80 and remove entries with NA for emotion scores
data_80img <- subset(data_img, emotion_score >= 0.80)
```
Inspect the data
```{r}
summary(data_80img)
```
Lets create a model that uses dominant_emotion, party_mentioned to predict ownership with nmo_title as a random intercept.
```{r}
mixed_model_img <- mblogit(ownership ~ party_mentioned + dominant_emotion, random = ~ 1 | nmo_title, data = data_80img)
```
The model does not converge.

Lets scrap this and try again with the happy_binary data.

Fit the mixed-effects multinomial logistic regression model:
```{r}
mixed_model_img <- mblogit(ownership ~ party_mentioned + happy_bin, random = ~ 1 | nmo_title, data = data_img)
summary(mixed_model_img)
```

Again, compare to the fixed-effects model for AIC and BIC scores.

```{r}
fixed_model_img <- mblogit(ownership ~ party_mentioned + dominant_emotion, data = data_img)
```

```{r}
AIC_fixed_img <- AIC(fixed_model_img)
BIC_fixed_img <- BIC(fixed_model_img)
AIC_mixed_img <- AIC(mixed_model_img)
BIC_mixed_img <- BIC(mixed_model_img)

cat("AIC Fixed:", AIC_fixed_img, "\n")
cat("AIC Mixed:", AIC_mixed_img, "\n")
cat("BIC Fixed:", BIC_fixed_img, "\n")
cat("BIC Mixed:", BIC_mixed_img, "\n")
```
```{r}
# Reorder the organisations so they appear nicely in the below graph
facet_order <- c("7 West Media", "Independent", "Australian Public Broadcaster", "Australian Community Media", "News Corp", "Nine Entertainment")
data_img$ownership <- factor(data_img$ownership, levels = facet_order)

# Combine party_mentioned and happy_bin into a single factor
data_img$party_happy <- interaction(data_img$party_mentioned, data_img$happy_bin, sep = "_")

# Create a custom order for the party_happy factor
custom_order <- c("alp_0", "alp_1", "coa_0", "coa_1")
data_img$party_happy <- factor(data_img$party_happy, levels = custom_order)

# Create the bar chart
bar_chart <- ggplot(data_img, aes(x = ownership, fill = party_happy)) +
  geom_bar(position = "dodge") +
  labs(x = "Ownership", y = "Count", fill = "Combination of \nPolitical Party\n and Happy Binary") +
  theme_minimal() +
  facet_wrap(~ ownership, scales = "free_x", strip.position = "bottom") +
  scale_fill_manual(values = c("alp_0" = "#d34547", "alp_1" = "#ff8b8f", "coa_0" = "#2a4e96", "coa_1" = "#92a6ce"),
                    labels = c("Labor + non-happy", "Labor + happy", "The Coalition + non-happy", "The Coalition + happy")) +
  theme(strip.text.x = element_blank())  # Add this line to remove x-axis labels on individual facets

# Print the bar chart
bar_chart

ggsave(filename = "happy_dist.pdf", plot = bar_chart, device = "pdf", width = 8, height = 5, dpi = 600)
```
 
```{r}
sessionInfo()
```





